Tackling discrepancies in freshwater invertebrate trait databases and taxonomic resolution: Harmonising across continents and trait aggregation

### 2021-01-20 Paper/Draft_harmonization_paper.tex
21.01.: Small changes in Paper

The outlined difficulties when working with or synthesizing trait data motivated us to (1) investigate discrepancies in trait definition between trait databases from Europe, North America, Australia, and New Zealand. As a basis for further evaluating different trait aggregation methods and how harmonising definition discrepancies between grouping features can influence trait-environment relationships, we established 4 novel invertebrate trait datasets from these trait databases by harmonising 7 grouping features. We evaluate different trait aggregation methods by (2) comparing aggregated traits with assigned traits by experts at family-level for two of the established trait datasets. (3) we re-analysed data on the effect of anthropogenic salinisation on biological traits by \citet{szocs_effects_2014} using trait data from the established harmonised European trait datasets and aggregated traits. By comparison with the original analysis, we investigated how harmonising and aggregating trait data can modify the outcome of trait-environment relationships.

@@

To examine the influence of the taxonomic hierarchy and the trait variability on the outcome of the different trait aggregation methods we created three hypothetical examples of different taxonomic hierarchies (scenarios). 
1) A family with an equal number of genera and species (5 genera each with 5 species respectively), denoted as \textit{sim\_base}.
2) A family where one genus has a much larger number of species than the other 4 genera (1 genus with 13 species, 4 genera with 3 species respectively), denoted as \textit{sim\_extreme}. 
3) A family where all genera have a different number of species (8, 2, 7, 3, 5), denoted as \textit{sim\_variation}. In total every family consisted of 25 species. To each hypothetical taxonomic hierarchy a hypothetical grouping feature with 3 traits was assigned. The 25 affinities for each trait were simulated by sampling from a truncated normal distribution with a mean value of 0.5 and 5 levels of standard deviations (0.2, 0.4, 0.6, 0.8, and 1) respectively to simulate different levels of trait variability. The truncated normal distribution was bound to 0 and 1. Simulated trait affinities were converted to percentages, similar to the data processing of the trait databases. The sampling was repeated 100 times for each standard deviation. Hence, in total 100 datasets for each level of trait variability were simulated, resulting in 12500 simulated trait affinities for each simulated trait ($25$ species per family $* 5$ levels of trait variability $* 100$ replicates). These 12500 simulated affinities per trait were assigned to each scenario. The above mentioned trait aggregation methods were applied to each simulated dataset. 
% ...resulting in 2500 aggregated affinities per trait and scenario ($12500$ trait affinities per trait divided by $25$, the number of species and multiplied by $5$, the number of aggregation methods). 
The results were compared based on the produced range of aggregated trait affinities between levels of trait variability and scenario. We also compared for each simulated dataset the differences in trait affinities obtained by each aggregation method.

@@

We evaluate the simulation results based on the produced range of aggregated trait affinities and by comparing results of every simulated dataset.

@@

While trait information from Europe and New Zealand was mainly on species-level, considerable portions of trait data from North America and Australia were on genus or family-level (Table \ref{tab:tax_coverage}). Assigned traits on family-level might not reflect the real trait diversity within a taxon, for example for Chironomidae (\cite{serra_synthesising_2016}). However, species identification can be complicated and time-consuming (\cite{marshall_taxonomic_2006, resh_which_2008}), and therefore trait information on family-level has been widely used, e.g. in bioassessment (\cite{beketov_spear_2009}).

>>>>

The outlined difficulties when working with or synthesizing trait data motivated us to (1) investigate discrepancies in trait definition between trait databases from Europe, North America, Australia, and New Zealand. As a basis for further evaluating different trait aggregation methods and how harmonising definition discrepancies between grouping features can influence trait-environment relationships, we established 4 novel invertebrate trait datasets from these trait databases by harmonising 7 grouping features. We evaluate different trait aggregation methods by (2) comparing aggregated traits with assigned traits by experts at family-level for two of the established trait datasets. (3) we re-analysed data on the effect of anthropogenic salinisation on biological traits by \citet{szocs_effects_2014} using trait data from the established harmonised European trait dataset and aggregated traits. By comparison with the original analysis, we investigated how harmonising and aggregating trait data can modify the outcome of trait-environment relationships.

@@

To examine the influence of the taxonomic hierarchy and the trait variability on the outcome of the different trait aggregation methods we created three hypothetical examples of different taxonomic hierarchies (scenarios). 
1) A family with an equal number of genera and species (5 genera each with 5 species respectively), denoted as \textit{sim\_base}.
2) A family where one genus has a much larger number of species than the other 4 genera (1 genus with 13 species, 4 genera with 3 species respectively), denoted as \textit{sim\_extreme}. 
3) A family where all genera have a different number of species (8, 2, 7, 3, 5), denoted as \textit{sim\_variation}. In total every family consisted of 25 species. To each scenario a hypothetical grouping feature with 3 traits was assigned. The 25 affinities for each trait were simulated by sampling from a truncated normal distribution with a mean value of 0.5 and 5 levels of standard deviations (0.2, 0.4, 0.6, 0.8, and 1) respectively to simulate different levels of trait variability. The truncated normal distribution was bound to 0 and 1. Simulated trait affinities were converted to percentages, similar to the data processing of the trait databases. The sampling was repeated 100 times for each standard deviation. Hence, in total 100 datasets for each level of trait variability were simulated, resulting in 12500 simulated trait affinities for each simulated trait ($25$ species per family $* 5$ levels of trait variability $* 100$ replicates). These 12500 simulated affinities per trait were assigned to each scenario. The above mentioned trait aggregation methods were applied to each simulated dataset. 
% ...resulting in 2500 aggregated affinities per trait and scenario ($12500$ trait affinities per trait divided by $25$, the number of species and multiplied by $5$, the number of aggregation methods). 
The results were compared based on the produced range of aggregated trait affinities between levels of trait variability and scenario. We also compared for each simulated dataset the differences in trait affinities obtained by each aggregation method.

@@

We evaluated the simulation results based on the produced range of aggregated trait affinities and by comparing results of every simulated dataset.

@@

While trait information from Europe and New Zealand was mainly on species-level, considerable portions of trait data from North America and Australia were on genus or family-level. Assigned traits on family-level might not reflect the real trait diversity within a taxon, for example for Chironomidae (\cite{serra_synthesising_2016}). However, species identification can be complicated and time-consuming (\cite{marshall_taxonomic_2006, resh_which_2008}), and therefore trait information on family-level has been widely used, e.g. in bioassessment (\cite{beketov_spear_2009}).
